---
title: "Hints for HW2:  data-storyteller"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, results ='show',include=TRUE,messages=FALSE)

library(tidyverse)
library(forcats)
library(ggplot2)
library(dplyr)
library(stargazer)
library(caret)
library(modelr)
```

# Reading in the data.
```{r getData, include=TRUE}
#replacing any possible blank entries as 'NA'.

storyteller<- read_csv("data-storyteller.csv", na = c(""))

#checking data types to see what may need changing
str(storyteller)
```


# Data Cleaning
Note: The School column is of the character type and it should be a factor. Other notes below in comments.

```{r cleanData1, include=TRUE}
storyteller$School<-factor(storyteller$School)

#The section column is of the Numeric type and should be be a factor instead

storyteller$Section<-factor(storyteller$Section)

#Each of the remaining columns is a discrete count of the students in each category.
#As it is not continuous the columns 'Very Ahead', 'Middling', 'Behind', 'More behind', 'Very behind'
#and 'Completed' should all be integers.

storyteller$`Very Ahead +5`<-as.integer(storyteller$`Very Ahead +5`)
storyteller$`Middling +0`<-as.integer(storyteller$`Middling +0`)
storyteller$`Behind -1-5`<-as.integer(storyteller$`Behind -1-5`)
storyteller$`More Behind -6-10`<-as.integer(storyteller$`More Behind -6-10`)
storyteller$`Very Behind -11`<-as.integer(storyteller$`Very Behind -11`)
storyteller$Completed<-as.integer(storyteller$Completed)

```

## Organizing the Data Structure

```{r Organize}

#Reordering columns to get a cleaner picture. I.E. 'Completed' being shifted to the other side.
#And section being a unique identifier is moved to the leftmost column.

storytellerTemp<-storyteller[,c(2,1,8,3,4,5,6,7)]

storyteller<-storytellerTemp

#displaying top 5 rows
head(storyteller)

```

## Missing Data

```{r cleanData2, include=TRUE}

#Checking for any NA values
sum(is.na(storyteller))

#There are no NA values in this dataset.

#The dataset is cleaned
head(storyteller)

```

# EDA and Data Viz

Observation: To be considered 'ahead' in any way, there are two categories: 'Very Ahead' and 'Completed'. There are 3 cateories in place to describe 'behind' which may lead the responses to be lopsided in favor of generally 'behind' over 'generally ahead'.

What other observations have you made?? Note them and clean the data appropriately. 



```{r viz, include=TRUE}

# Creating a bar chart to show the number of sections from each school
SchoolValues<-c(length(which(storyteller$School=='A')), length(which(storyteller$School=='B')), length(which(storyteller$School=='C')), length(which(storyteller$School=='D')), length(which(storyteller$School=='E')))

barplot(SchoolValues, names.arg = c('A', 'B', 'C', 'D', 'E'), main='Number of sections Per School')

#plotting section and Completed and summarizing the data

barplot(storyteller$Completed, main='#completed students / section', names.arg = c(1:30))
summary(storyteller$Completed)
```

## Further Data Cleaning based on EDA and Viz

```{r cleanData3, include=TRUE}

#Plotting section and Very Ahead and summarizing

barplot(storyteller$`Very Ahead +5`, main='number of very ahead students per section', names.arg = c(1:30))

summary(storyteller$`Very Ahead +5`)
#plotting section and Middling and summarizing

barplot(storyteller$`Middling +0`, main='number of Middling students per section', names.arg = c(1:30))

summary(storyteller$`Middling +0`)
#plotting section and Behind and summarizing

barplot(storyteller$`Behind -1-5`, main='number of Behind students per section', names.arg = c(1:30))

summary(storyteller$`Behind -1-5`)
#plotting section and More Behind

barplot(storyteller$`More Behind -6-10`, main='number of More Behind students per section', names.arg = c(1:30))


summary(storyteller$`More Behind -6-10`)
#Plotting section and Very Behind

barplot(storyteller$`Very Behind -11`, main='number of Very Behind students per section', names.arg = c(1:30))

summary(storyteller$`Very Behind -11`)
#determining the number of students in each category
StudentSums<-colSums(storyteller[,3:8])

sum(StudentSums)

#determining the amount of students in each section

SectionSums<-rowSums(storyteller[,3:8])

data.frame(SectionSums)

#Creating a barplot to show distribution

StudentSums<-colSums(storyteller[,3:8])
barplot(StudentSums, main="Student totals across all categories")

```

Note: There is a gap between middling and completed, meaning that the only students who can be described #as 'ahead' have already finished the program.

There is no category that describes between 'Middling' and 'Very Ahead'. This means that survey respondents would have to decide between 'middling' or 'Very Ahead'; should they fall in the middle? This may skew the data towards one side or the other, however (Assumption) it
is reasonable to assume that survey respondents would stick to a more honest 'Middling Answer'
rather than exaggerating their students' success.

Further Note: The data seems to be centered on the 'Behind' category, and by a large margin.

## EDA (cont.)

```{r EDA, include=TRUE}

StudentSums/sum(StudentSums)

```

More observations:
 - 14% of students are on track in the middling category.
 - Nearly 20% of students have completed the program.
 - Nearly 50% of students in this program are in the 'behind' category alone.
 - Students in the bottom two categories make up 20% of the sample. Meaning ~70% of students are behind in the curicculum. 

```{r middling, include=TRUE}

plot(storyteller$Completed, storyteller$`Middling +0`)

```

If we were to assume that students are generally not doing well because the program is difficult,  #there is is still a discrepancy with the number of students that have 
completed the course ahead of schedule. Essentially, A fifth of the students have nothing to
work on for the remainder of the time.

So, lets break the data down into individual schools to see if we can better understand the data.

```{r groupAnalysis, include=TRUE}

storytellerA<-storyteller[which(storyteller$School == "A"),]
storytellerB<-storyteller[which(storyteller$School == "B"),]
storytellerC<-storyteller[which(storyteller$School == "C"),]
storytellerD<-storyteller[which(storyteller$School == "D"),]
storytellerE<-storyteller[which(storyteller$School == "E"),]

StudentSumsA<-colSums(storytellerA[3:8])
StudentSumsA

barplot(StudentSumsA, main = "School A")

StudentSumsB<-colSums(storytellerB[3:8])
StudentSumsB

barplot(StudentSumsB, main = "School B")

StudentSumsC<-colSums(storytellerC[3:8])
StudentSumsC

barplot(StudentSumsC, main = "School C")

StudentSumsD<-colSums(storytellerD[3:8])
StudentSumsD

barplot(StudentSumsD, main = "School D")

StudentSumsE<-colSums(storytellerE[3:8])
StudentSumsE

barplot(StudentSumsE, main = "School E")

```

# Initial Observations and Remarks

Observe:

 - Schools B and D do not follow the same picture that was shown by the data combined. 
 - A, C, and E follow generally the same pattern as shown by figure 1
 - B shows many students ahead of the curriculum, completing the program. 
 - The majority of students that are behind are behind by 1-5 assignments and very few are in the lowest two categories. 
 - This is the most positive picture that the data shows out of the schools.
 - School D is the opposite with a small percentage of students having completed all assignments and a large percentage of students being considered 'Very Behind'. 

Comparing the two schools as a representation of the population of program takers ... 

```{r cleanData, include=TRUE}
sum(StudentSumsB)/sum(StudentSums)

sum(StudentSumsD)/sum(StudentSums)
```

School D has a small number of students, meaning that the data from school D has a greater probability of not being representative of the situation and may be acconted for by random chance. School B is a sample that is approximately one quarter of the overall population, making it more representative of the situation. It also shows that school B is doing something that the rest of the schools are not in order to get students through the course.

```{r finalANA, include=TRUE}

#looking at the data to determine quality sections

barplot(storytellerB$Completed, names.arg =c(1:12),ylim=c(0,30), main = "Students Completed by Section")
```
Sections 6, 10, and 12 have the most promising results within school B.

```{r final2, include=TRUE}
rowSums(storytellerB[,3:8])

barplot(storytellerB$`Behind -1-5`+storytellerB$`More Behind -6-10`+storytellerB$`Very Behind -11`, names.arg =c(1:12), ylim=c(0,30), main = "Students Behind by Section")
```
